%
% file: dedication.tex
% author: V?ctor Bre?a-Medina
% description: Contains the text for thesis dedication
%

\chapter*{Dedication and acknowledgments}
\begin{SingleSpace}
\section*{Acknowledgments}
First and foremost, I would like to thank my both supervisors Prof. Sergio Terzi from Politecnico Di Milano and Dr. Luigi Serio at European Council for Nuclear Research CERN. 

I am grateful to Professor Sergio Terzi who agreed to supervise and guide me throughout. This accomplishment would not have been possible without his academic and moral support. I am also grateful with the bottom of my heart for the opportunity to be a part of Dr.Luigi Serio administration, resources and performance group in engineering department at CERN. Taking this as an opportunity, I want to show my profound gratitude to everyone who has facilitated me to fulfill this task within the deadlines. 

Most importantly, I want to acknowledge European Council for Nuclear Research and especially my CERN supervisor, Dr. Luigi Serio.  I would like to thank him and share my deepest gratitude for the time he took to discuss my work, providing me with deep in insights into the Project management for complex projects like nuclear infrastructure. He pushed me to work independently and always provided his counsel whenever I needed it. I have learned a great deal working with him and always got inspired by his own example to work hard and expand my knowledge and thinking. His optimism and kindness motivates me to keep searching for greater heights.

My thesis would not have been possible without the amiable support of two of my good friends Gul Muhammad and Keshava Prasad Gubbi and for their fruitful discussions about writing and organizing my thesis. 

I am eternally grateful to my Father and Siblings who have always encouraged me to pursue my dreams. They have always believed in me and are my pillars of support.

I would like to dedicate this work to my life-coach my Late mother Bimla Kandhol, who had always encouraged me to achieve growth and success in life. I always miss your voice and presence. 

\end{SingleSpace}
\clearpage